\documentclass[11pt]{article}

\usepackage[margin=0.8in]{geometry}
\usepackage{setspace}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{array}
\usepackage{ragged2e}

\definecolor{PrimaryRed}{HTML}{8B1E1E}
\definecolor{SoftBlue}{HTML}{DDEEFF}
\definecolor{DarkGray}{HTML}{333333}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=black,
    citecolor=black
}

\titleformat{\section}{\large\bfseries\color{PrimaryRed}}{}{0em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

\setstretch{1.15}

\begin{document}

\begin{titlepage}
\newgeometry{margin=0.85in}
\thispagestyle{empty}

\begin{center}
    \vspace*{0.25in}
    {\large University of Rochester\\}
    {\normalsize Warner School of Education\\[0.35in]}

    {\color{PrimaryRed}\rule{\textwidth}{1.4pt}}\\[0.22in]
    {\Huge\bfseries Joy-Oriented Literacy Pursuit\\[0.12in]}
    {\Large Formal Reflective Journal\\[0.22in]}
    {\color{PrimaryRed}\rule{0.72\textwidth}{0.7pt}}\\[0.32in]

    \colorbox{SoftBlue}{%
        \parbox{0.86\textwidth}{%
            \centering\vspace{0.12in}
            \textbf{Classroom Technology, Design, and Literacy as Joy}\\[0.05in]
            A Teaching Neurodivergent Journal focused on Literature \& Joy via a Puzzle Plan
            % A Journal Kindergarten circuits, Grade 4 Minecraft, and Grade 5 airplane/fan-circuit design
            \vspace{0.12in}
        }%
    }
\end{center}

\vspace{0.75in}

\begin{center}
\renewcommand{\arraystretch}{1.35}
\begin{tabular}{>{\bfseries\RaggedLeft}p{1.55in} p{4.85in}}
Student & Piter Garcia \\
Course & EDU498: Culturally-Historically Responsive Literacies \\
Class Session & Module 3 / Week 5 \\
Assignment & Joy-Oriented Literacy Pursuit \\
Context & Teaching Reflection through the Lens of this course \& a Puzzle Plan\\
Submission Date & June 05, 2026 \\
\end{tabular}
\end{center}

\vfill

\begin{center}
\begin{minipage}{0.78\textwidth}
\centering
\itshape
Joy-oriented literacy is not extra. It is part of access. When students feel capable, connected, and allowed to participate as themselves, literacy becomes a tool for liberation, self-determination, and empowerment.
\end{minipage}
\end{center}

\restoregeometry
\end{titlepage}
\clearpage

\section*{What I Created and Who I Shared It With}

For my joy-oriented literacy pursuit, I created and facilitated classroom-based technology and design lessons where students experienced literacy through building, collaboration, speaking, movement, problem-solving, and shared excitement. I focused on moments where students could use language, materials, technology, and relationships to make meaning aligned with the lesson goals. The examples that brought the most joy came from my Kindergarten airplane, light-bulb, and fan squishy circuit lessons, as well as my Minecraft lessons in third (3) and fourth (4) grades. These activities were shared with students and with the classroom community itself because students built, watched, helped, responded, revised, and learned from one another in real time.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/joy-evidence-from-placement.md}{\textit{Placement evidence: six documented examples across Grade 4 and Kindergarten}}}

In my Kindergarten circuit lessons, I gathered students around a demonstration where they learned about batteries, switches, wires, Play-Doh, and motor-based fans by watching, naming parts, holding materials, connecting pieces, and counting down together before turning on a built motor-based fan. These lessons became shared literacy events because students participated physically, verbally, visually, and socially. In the airplane-building squishy circuit lesson, students worked in teams to build with Play-Doh and motor-based fans, taking roles such as building wings, shaping the body, making wheels, and connecting the fan to a switch.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/kindergarten-circuit-demo-airplane-prep-evidence.md}{\textit{Kindergarten transcript evidence — circuit demo and airplane prep}}} In my fourth (4) grade Minecraft lessons, students who finished their work helped classmates who were not done, turning digital literacy into peer teaching and community responsibility.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/grade-4-minecraft-agent-challenge-evidence.md}{\textit{Grade 4 transcript evidence — Minecraft agent challenge}}} These moments showed that a shared experience can be a literacy product, where students become makers, helpers, designers, and explainers.

\section*{My Creative and Learning Process}

My creative process began with a question: \textit{how can technology lessons become more than task completion?} Knobel and Lankshear helped me understand this as a question about new literacies, where meaning-making happens through digital, visual, participatory, and networked practices. Price-Dennis, Holmes, and Smith helped me see the same question as one of access: digital tools matter when they make curriculum more reachable and allow students multiple ways to participate. Brooks and Frankel also reminded me that students do not enter classrooms as blank technology users; they bring languages, cultures, family histories, interests, identities, and community knowledge with them. Together, these readings helped me view my lesson activities as literacy work because students were interpreting systems, using vocabulary, communicating with peers, reading visual and technical cues, and representing their thinking through more than one mode.

In the light-bulb and airplane building squishy lessons, I used switches, batteries, wires, Play-Doh, fans, and model-building pieces to connect vocabulary to action. Students did not only hear the word ``switch''; they saw it, held it, connected it, and watched what happened when it worked.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/kindergarten-circuit-demo-airplane-prep-evidence.md}{\textit{Evidence: Example 5 — vocabulary and collective participation}}} Students also decided who would make each part of the airplane, which made the activity collaborative and distributed. Similarly, in one fourth (4) grade Minecraft lesson, I treated students who finished early as class resources rather than treating completion as the end of learning. That shifted the process from ``I am done'' to ``I can help someone else succeed,'' helping one student navigate and troubleshoot while the other student deepened mastery by explaining.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/grade-4-minecraft-agent-challenge-evidence.md}{\textit{Evidence: Example 2 — peer teaching as joyful expertise}}}

This process taught me to notice joy as part of learning design. Joy did not happen only because the activities were fun. It happened because students had access to materials, roles, language, peer support, and visible success. Pinkard's ecosystem view of digital literacy helped me name this more clearly: students need tools, mentors, audiences, feedback, and real reasons to create. Stornaiuolo and Thomas also helped me see digital literacy as agency, where young people use tools to participate, speak, and organize meaning. That became the design principle behind this pursuit: \textit{students needed enough structure to feel safe and enough freedom to feel ownership}.

\section*{How This Cultivated Joy, Love, and Aesthetic Fulfillment}

This literacy pursuit cultivated joy because students were able to see their learning become real. In my Kindergarten airplane-building and fan squishy circuit lessons, the joy came from collective anticipation. Students watched the pieces connect, counted down together, and saw the fan turn on.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/kindergarten-circuit-demo-airplane-prep-evidence.md}{\textit{Evidence: Example 5 — collective countdown and fan circuit demo}}} That moment mattered because the science was no longer abstract; it became visible, audible, physical, and shared. The joy was not only in the fan working; it was in the class experiencing success together.

These lessons also created aesthetic fulfillment because students built something with shape, motion, design, and purpose. Their work had form, creativity, visible effort, and a relationship to the physical world. The pursuit also cultivated love through relationships. In my fourth (4) grade Minecraft lessons, students who already understood the task helped classmates who were still working. That peer support changed the emotional meaning of the activity. Success was not private or competitive; it became something students could give to each other.\footnote{\href{https://github.com/pzg8794/EDU498-joy_oriented_literacy_pursuit_journal/blob/main/grade-4-minecraft-agent-challenge-evidence.md}{\textit{Evidence: Example 2 — peer teaching in the Minecraft session}}}

These activities reminded me that joy in literacy does not always look quiet or neat. Sometimes joy looks like a student trying again, a group arguing productively over what goes where, a fan finally turning on, a peer stepping in to help, or a student realizing that their hands, ideas, and voice matter. For me, that is aesthetic fulfillment: the moment when learning feels alive, when students can see themselves inside the work, and when the classroom becomes a place where knowledge is built rather than simply delivered.

\section*{Muhammad's Literacy Pursuit and Digital Literacy Readings}

\textit{Muhammad's framework} helps me understand that a literacy pursuit should develop identity, skills, intellect, criticality, and joy together. This project was not designed only to teach science skills. It also helped students develop vocabulary and technical understanding while seeing themselves as builders, helpers, problem-solvers, and contributors. That matters because literacy should help students understand themselves, participate in community, and experience agency.

Knobel and Lankshear strengthened this connection by helping me name these activities as new literacies: meaning-making can happen through digital, visual, participatory, and networked forms. Price-Dennis, Holmes, and Smith connect this work to inclusive digital literacy because students need more than one pathway to demonstrate understanding. Brooks and Frankel's work on transnational digital literacy also matters because students bring languages, cultures, family histories, interests, identities, and community knowledge with them. When a student builds in Minecraft, explains a circuit orally, draws a design, or helps a peer troubleshoot, that student is using literacy in a way that deserves recognition.

Melo-Pfeifer and Gertz pushed me to think about criticality because digital and media spaces are not neutral. Students need to question headlines, images, missing voices, representation, credibility, and assumptions. The Common Sense Digital Citizenship materials extend this by giving students language for privacy, audience, confirmation bias, online identity, community responsibility, and upstander communication. My \textit{Puzzle Plan} focuses on building the habits that critical digital literacy needs: collaboration, questioning, explanation, audience awareness, and responsibility to others.

Stornaiuolo and Thomas helped me understand agency by showing that young people can use digital tools to speak, organize, and participate in public life. Pinkard's ecosystem view of digital literacy reminds me that students need tools, mentors, audiences, feedback, and real reasons to create. Serpagli and Mensah's work with social media-style science learning connects to my placement examples because students were using active and shared forms to explain technical ideas. Together, these readings show that digital literacy should position students as creators, not just consumers.

\section*{How This Informs My Future Teaching}

This experience reminds me that literacy is not limited to written language. Literacy is also building, coding, designing, explaining, translating, collaborating, troubleshooting, helping, evaluating media, and trying again. In my classroom, I will design activities where students have multiple entry points and different ways to show their understanding. This is especially important for students who may be multilingual, neurodivergent, disabled, chronically ill, culturally diverse, or simply different from the ``default'' learner that schools often imagine.

In practice, I will continue designing projects where students create something meaningful and share it with an authentic audience. Students could build Minecraft worlds connected to a science concept, create bilingual or multimodal explanations, make collaborative classroom murals, create technology tutorials for peers, or build simple circuitry projects and explain how they work through speaking, drawing, writing, or video. I will also make peer teaching part of the classroom culture because students often understand each other's barriers in ways adults can miss.

My biggest takeaway is that joy-oriented literacy is part of access. Through my \textit{Puzzle Plan and EQUITAS lens}, the question is not whether a task looks traditional enough to count as literacy. The question is whether the task gives students real access to meaning, identity, criticality, collaboration, and agency. When students feel capable, connected, and allowed to participate as themselves, literacy becomes a tool for liberation, self-determination, and empowerment. This pursuit helped me see that my teaching practice should ask whether students had the chance to feel ownership, joy, relationship, and possibility.


\end{document}